Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-11-19 contents/draft.tex
Updates from Overleaf

\begin{document}
    \title{Improvement of Korean Morphological Analysis System \\Through Transformer-based Re-ranking}

@@

    \abstract[Abstract]{Korean morphological analysis plays a basic and important role enough to be called the first step in Korean language analysis. Due to the nature of Korean agglutinative words, it was difficult to build an automatic analysis system because the analysis was not completed with part-of-speech tagging alone, as in English. In addition, various methods for morphological analysis have been proposed, but efficient methods such as BPE are mainly used in applications intended to be used as tokenizers for deep learning. In this paper, we propose a method to maximize the performance of an efficient morphological analysis system that can be used for tokenization through multi-stage re-ranking based on deep learning. For a number of various cases whose rankings have been changed through re-ranking, it is possible to improve performance while maintaining speed by updating the cost matrix of the lattice-based morphological analysis system in the future. Through experiments, we showed that the proposed method effectively improved performance by reducing errors by more than 30\% in both the spoken language model and the written language model.}
    \keywords{Korean morphological analysis, natural language understanding, deep learing, pretrained transformer encoder, re-ranking}

@@

    Korean morphological analysis is the process of determining parts of speech by finding morphemes, which are the smallest units of linguistic expression that have an independent meaning in a sentence.
    In an isolating language like English, this can be done relatively easily by tagging parts of speech sequentially, but in Korean, the nature of the agglutinative language requires separating endings or investigatives and restoring inflections to their original form.
    In addition, since the basic input of other Korean analysis tasks is often a separated morpheme, the accuracy of morphological analysis greatly affects the performance of Korean analysis.
    Modern high-performance deep learning methods in natural language processing use a tokenization process that breaks text into smaller units, and then converts each token into a vector as an input to the computational model~\cite{Mikolov2013}.
    In this case, the token unit is mainly subword units, and to reflect the characteristics of Korean, subword tokenization with separated morphemes is attempted in advance~\cite{SongHJ2021}.
    Using the results of morphological analysis for this tokenization process improves the overall performance of the analysis by reflecting the semantic units of Korean, but requires a highly accurate and fast morphological analyzer.

    Various approaches have been proposed for morphological analysis, which plays a fundamental and important role in Korean language analysis~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    In general, when people understand speech or writing, they try to make sense of it using the vocabulary and concepts they know.
    While there are ways to use rules or dictionaries to reflect this way of understanding~\cite{KwonHC1991}, the problem is that it becomes difficult to build and maintain a dictionary for vocabulary that appears in every text.
    Therefore, methods for tagging syllable units without a dictionary have been proposed~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016}, and studies to improve them have been carried out continuously~\cite{KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    From a mechanical point of view, syllable-by-syllable morphological analysis can be done either by tagging syllable-by-syllable and then applying a base form restoration dictionary~\cite{ShimKS2011, LeeCH2016}, or by tagging syllable-by-syllable with the base form already restored~\cite{YounJY2021}.
    However, syllable-by-syllable morphological analysis has limitations in that it is difficult to accurately identify morpheme boundaries, and it is difficult to learn long-term contextual information as the length of the sequence increases.
    In this paper, the former is referred to as dictionary-based morphological analysis, and the latter is referred to as syllable-unit morphological analysis.
    Both methods also share the limitation that they are trained on a manually labeled corpus and cannot accurately analyze new syllable combinations or morphemes that do not appear in the training corpus.
    In recent years, with the development of the Internet and the spread of open source and open data, web texts, corpora, language resources, and knowledge shared by different people have accumulated significantly.
    The reduced cost of building and maintaining a dictionary can be a great opportunity to overcome the limitations of dictionary-based methods.

    Against this background, this paper considers how the dictionary-based morphological analysis method used by MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis in tokenizers, an essential preprocessing tool for deep learning, can be effectively improved by deep learning, and proposes a method.
    The dictionary-based morphological analysis method~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018} trained by the CRF method~\cite{Lafferty2001} lists the candidate morphemes in the dictionary from a given sentence to form a lattice structure connected by a directed graph, and finds the optimal morphological analysis path within it.
    The process of finding the best path in the lattice uses the Viterbi algorithm~\cite{Viterbi1967}, which finds the path that minimizes the cost of each morpheme node and the sum of the neighborhood costs of two consecutive morphemes.
    The main types of errors in these dictionary-based morphological analysis methods are when new words not in the dictionary are used in a sentence, or when the optimal path calculation selects the wrong result due to a bias.
    For example, it may be less costly to select one long morpheme than several short morphemes, but it may sometimes be the wrong analysis.
    The main motivation for this research is that the path that minimizes the sum of all costs for nodes and links may not be the optimal path.

    In order to identify the cases where a suboptimal solution is actually the best solution according to the best path calculation, we modified the best path calculation method to generate suboptimal analysis results and check the extent to which they are correct.
    While there are many different ways to choose the next best path, we used the method of replacing a morpheme node on the optimal path with a lower ranked node.
    As shown in Table~\ref{tab:maximum-performance}, we were able to confirm the extent to which analysis performance could be improved by replacing the optimal path with a lower-ranked node.
    We can think of the problem of finding the actual correct answer among these generated suboptimals as similar to the problem of re-ranking search results in information retrieval~\cite{BaeYJ2021}.
    Previously, in ~\cite{ChoiYS2018}, the N-best analysis results generated by the seq2seq model were re-ranked based on a convolutional neural network to improve performance.
    In this study, the re-ranking was performed using two BERT models of different types and forms as proposed in ~\cite{Nogueira2019}.
    The experimental results show that the first stage re-ranking improves the performance by more than 20\% over the previous written and spoken models, and the second stage re-ranking with a different type of input and a different type of pre-trained model further improves the performance by more than 30\% over the previous written and spoken models.

    With this method, the performance of the dictionary-based morphological analysis method could be further improved, but the overall analysis time increases when the morphological analysis system is configured including the re-ranking model itself.
    However, it is possible to use the results of multiple re-ranked morpheme analyses to update the connection costs between morphemes in the dictionary, similar to the backpropagation process in a typical neural network.
    It is also expected that the morphological analysis system with improved connection costs will be able to generate better re-ranking candidates, which will further improve performance by doing so iteratively.
    Further research is needed for this in the future, and in this paper, only the performance improvement through the second stage re-ranking was covered as the scope of the study.
    The main contributions of this study are as follows:
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We explore the possibility of performance improvement by introducing a method to replace the optimal path with a suboptimal node, and propose a method to effectively improve the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To improve the performance of dictionary-based analysis by re-ranking the morphological analysis results, we propose to extend the performance improvement by using different BERT models to perform two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We propose a new method for updating connection costs in dictionaries based on the results of reranked morphological analysis, and provide directions for future research by suggesting improvements and directions for future research.
    \end{enumerate}
    These contributions provide important insights into the performance improvement of Korean morphological analysis and the direction of future research, and will serve as a useful reference for future researchers.

    This paper is organized as follows:
    In Section ~\ref{sec:related-work}, we introduce previous research cases related to this study, and in Section ~\ref{sec:morphological-analysis-model}, we discuss how to configure and train a dictionary-based morphological analysis system.
    In Section ~\ref{sec:re-ranking-model}, we discuss how to generate secondary results of morphological analysis and produce re-ranking data, and propose a method for training a two-stage re-ranking model.
    In Section ~\ref{sec:results}, we discuss the results of performance improvement by morphological analysis model and re-ranking model.
    Finally, in Section ~\ref{sec:conclusion}, we conclude the paper and discuss the limitations of this study and the direction of future research.

@@

    Korean morphological analysis has seen an influx of various methodologies over the years~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}. The nature of the Korean language, being agglutinative, introduces challenges that have propelled researchers to come up with inventive solutions, many of which have laid the groundwork for future research.

    \subsection{Traditional Dictionary-based Approaches}\label{subsec2.1}
    The earliest attempts at Korean morphological analysis heavily depended on rule-based systems and dictionary-based methods~\cite{KwonHC1991}. These methodologies essentially employed predefined sets of linguistic rules or large dictionaries to detect morphemes and determine parts of speech. One of the most significant advantages of such an approach is its deterministic nature, which can lead to high accuracy when the input text closely adheres to the rules or dictionaries used. However, they are notably challenging to scale and update, particularly with the constant evolution of the language and emergence of new words. The dynamic nature of language, especially in the age of the internet, has made maintaining comprehensive dictionaries labor-intensive.

    \subsection{Syllable-unit Morphological Analysis}\label{subsec2.2}
    In an attempt to bypass the limitations of dictionary-dependence, syllable-by-syllable morphological analysis emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}. This method either tags each syllable and then applies a base form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}. One notable drawback is the challenge in accurately pinpointing morpheme boundaries. Furthermore, as sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Deep Learning and Tokenization}\label{subsec2.3}
    The advent of deep learning has reshaped the landscape of Korean morphological analysis. High-performance deep learning models in NLP now employ tokenization, dividing the text into smaller units and then converting these tokens into vectors as an input to the computational model~\cite{Mikolov2013}. Especially significant is the attempt to conduct subword tokenization, incorporating separated morphemes to cater to Korean linguistic characteristics. In particular, it is noteworthy that we attempted to tokenize subwords by incorporating segmented morphemes to suit the linguistic characteristics of Korean, and it is known that incorporating accurate morpheme analysis results into tokenization can improve overall analysis performance.

    However, with the progression of technology and accessibility to vast amounts of data, the possibility of integrating dictionary-based methods with deep learning techniques emerged. As motivated in our study, MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis, utilized dictionary-based morphological analysis trained by the CRF method~\cite{Lafferty2001}. This method attempts to find the optimal morphological analysis path by forming a lattice structure as in ~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018}. The amalgamation of dictionary-based methods with deep learning, as explored in our research, signifies the latest stride in this journey, aiming to harness the strengths of both approaches. Future research avenues include refining these hybrid methods and further exploring the capabilities of end-to-end models for this complex task.

@@

    A representative corpus for training Korean morphological analysis models is the Sejong Morphological Analysis Corpus, often referred to as the Sejong Corpus. The Sejong Corpus, a major output of the 21st Century Sejong Project, contains 15 million words and remains a mainstay of Korean morphological analysis research~\cite{ChoeMW2008}.

    Subsequent efforts have extended and refined this foundational work. The University of Ulsan, addressing the limitations of the Sejong corpus~\cite{KimIH2010}, introduced [UCorpus], an expanded corpus of 63 million words. This expanded corpus not only added new data, but also corrected errors in the original Sejong corpus~\cite{UCorpusHG}. To take this a step further, the National Institute of the Korean Language launched a massive Korean corpus construction project in 2018. The fruits of this labor, a corpus called <Everyone Corpus>, were released to the public in 2020~\cite{EveryoneCorpus}. A special offshoot of this project produced a morpheme-tagged corpus of 3 million words, including spoken language~\cite{KimIH2019}.

@@

    In order to effectively train a dictionary-based morpheme analysis model, the morpheme-tagged corpus, typically represented in lemma form, needs to be transformed to include boundary information between morphemes in its surface form. Crucial to this transformation is string alignment, a process that accounts for discrepancies between lemma and surface forms in the Korean morphological analysis corpus.

    In this research, string alignment was performed using the Smith-Waterman algorithm, which uses a scoring matrix based on the similarity of the grapheme unit of Korean letters for each word pair (as shown in Figure~\ref{fig:sample}). Each aligned sentence containing a morpheme tag can be converted into a training sample tailored for dictionary-based morphological analysis.

    The resulting table data in Figure~\ref{fig:sample} illustrates this process. Here, each row acts as a lexical unit. The first four columns contribute to feature generation, while the last four columns facilitate post-lemmatization. Using the above morphological corpus, a large number of training samples can be generated according to the process shown in Figure~\ref{fig:sample}. Except for the evaluation samples, the remaining sentences are used to train the dictionary-based morphological analysis model via the CRF algorithm. The output of this training allows the calculation of the costs associated with each morpheme node and the linking of two consecutive morphemes. This in turn allows the discovery of the optimal path using the Viterbi algorithm.

@@

    Figure~\ref{fig:lattice} shows a snapshot of the lattice structure, an integral part of morphological analysis. Part (1) shows a fragment of the lattice structure formed when the example sentence from Figure~\ref{fig:sample} is entered. Meanwhile, part (2) shows the optimal path as determined by the Viterbi algorithm.

    However, the path inferred by the trained model may differ from the correct solution constructed by a human. The nodes marked with stars in part (1) represent the correct nodes. The upper left number of each node indicates the ranking of the nodes accessible at each decoding point. It can be seen that the choices made at certain moments can deviate from the correct solution. In order to improve the performance of the analysis, it is imperative to develop mechanisms to correct these discrepancies.

@@

    While dictionary-based morphological analysis offers significant advances in Korean language processing, its optimal paths occasionally diverge from the correct solutions that humans understand. This divergence underscores the need for a model that re-evaluates these primary results and reorients them to achieve higher accuracy. This method, called "re-ranking," involves generating multiple analyses of an input and then reordering them based on a new set of criteria or models, thereby improving the overall quality of the results.

@@

    Before re-ranking can begin, multiple analyses, typically referred to as N-best paths, of the input sentence are generated.
    This involves extracting the top N analyses from the lattice structure in general.
    In this study, a novel approach has been introduced to generate secondary paths, as shown in part (3) of Figure~\ref{fig:lattice}, by selecting the second-best node instead of each best node constituting the path from the best-path result.
    Encouragingly, some of these secondary paths provided alternatives that reconciled incorrect answers with the correct ones.
    Similarly, paths modified by favoring the third-best node were called tertiary paths, and this naming convention continued for subsequent paths.
    In our preliminary test, the secondary paths, including the optimal and some suboptimal paths, were shown to cover the majority of correct morphological analyses as measured by human judgments. (See Table~\ref{tab:maximum-performance})

@@

    BERT (Bidirectional Encoder to Transformer Representation) models~\cite{Devlin2019} have revolutionized many natural language processing tasks by understanding the context in which words appear in text. In this study, we attempt to leverage the power of BERT to reorder the generated secondary paths. We labeled the generated secondary paths with scores related to morphological analysis performance in order to fine-tune a pre-trained BERT model specialized for Korean with a large amount of Korean text. After pre-testing several scoring methods on a modest scale, we found that using scores based on the degree of error, rather than scores based on accuracy, to widen the gap between correct and incorrect answers is effective for learning.

    Once the BERT model has been fine-tuned and trained for the reranking task, it will be able to predict a reranking score for each path in the secondary path list. This means that, taking into account the context, morphological organization, and other essential linguistic features of the path, the model will assign a score for each path. The paths are then re-ranked according to this score, and the path with the highest score is selected as the best morphological analysis result.

@@

    Given the complexity of the Korean language, a single reranking step does not always yield the most accurate results. Therefore, we propose a two-step reranking approach, as in ~\cite{Nogueira2019}.

    In the first step, we re-rank the secondary paths generated using the BERT model as described in section ~\ref{subsec4.3}. In the second step, we introduce another BERT variant that is optimized for a different set of linguistic features or trained on a different dataset. This allows us to perform a fine-grained re-evaluation, further refining the list and pushing the more contextually accurate paths to the top.

    As shown in Figure~\ref{fig:ranking}, for a two-stage reranking model, the first stage performs a first reranking, taking as input a secondary path in morphologically tagged lemma form. It then performs a second reranking, again taking as input the path reranked in stage 1 and the original input sentence. This is done to improve the effectiveness, since it didn't work very well when given the same type of input.

    \subsection{Iterative Updatability}\label{subsec4.5}

    The re-ranked paths are then integrated with the original dictionary-based morphological analysis. The connection costs between morphemes in the dictionary are updated based on feedback from the re-ranking model. Over time, this feedback loop ensures that the morphological analysis model becomes more closely aligned with human understanding and minimizes discrepancies in optimal path selection.

@@

    In our effort to enhance Korean morphological analysis, we evaluated the performance of the proposed deep learning integrated dictionary-based morphological analysis method. This section presents the results of the experimental evaluation, taking into account the improvement over conventional methods and the effectiveness of our re-ranking model.

@@

    We found that the proposed transformer-based reranking technique consistently improves the results of existing morphological analysis models. These results open up new possibilities in the area of Korean morphological analysis by further improving the results of existing traditional machine learning models.
    Finally, since the results of the major related works proposed in the literature are mostly conducted on the Sejong corpus, we would like to compare the performance improvement of the Sejong corpus with the results of previous studies.

>>>>

\begin{document}
    \title{Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking}

@@

    \abstract[Abstract]{Korean morphological analysis is a fundamental step in analyzing the Korean language. Owing to the nature of Korean agglutinative words, building an automatic analysis system is difficult because parts-of-speech tagging independently is insufficient to complete analysis, as in English. In addition, various methods for morphological analysis have been proposed; however, efficient methods, such as BPE (Byte Pair Encoding), are mainly used in applications intended to be used as tokenizers for deep learning. In this study, we propose a method to maximize the performance of an efficient morphological analysis system that can be used for tokenization through multi-stage re-ranking based on deep learning. For cases whose rankings have been changed through re-ranking, it is feasible to improve the performance while maintaining speed by updating the cost matrix of the lattice-based morphological analysis system. Through experiments, we demonstrate that the proposed method effectively improves performance via reducing errors by over 30\% in both spoken and written language models.}
    \keywords{Korean morphological analysis, natural language understanding, deep learning, pretrained transformer encoder, re-ranking}

@@

    Korean morphological analysis is the process of determining parts of speech by identifying morphemes, which are the smallest units of linguistic expression with independent meanings in a sentence.
    In an isolating language, such as English, this identification can be achieved relatively easily by tagging parts of speech sequentially. However, in Korean, the nature of the agglutinative language requires separating endings or postpositions and restoring inflections to their original form.
    In addition, because the basic input of other Korean analysis tasks is often a separate morpheme, the accuracy of the morphological analysis significantly affects the performance of Korean analysis.
    Modern high-performance deep learning methods in natural language processing (NLP) use a tokenization process that breaks the text into smaller units and converts each token into a vector as an input to the computational model~\cite{Mikolov2013}.
    Here, the token unit is mainly a subword unit, and to reflect the characteristics of a Korean subword, tokenization with separate morphemes is attempted in advance~\cite{SongHJ2021}.
    Using the results of the morphological analysis for this tokenization process improves the overall performance of the analysis by reflecting the semantic units of Korean. This requires a highly accurate and fast morphological analyzer.

    Various approaches have been proposed for morphological analysis, which is crucial in Korean language analysis~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    In general, when people understand speech or written text, they attempt to make sense of it using vocabulary and concepts that they are familiar with.
    While there are approaches to use rules or dictionaries to reflect this understanding~\cite{KwonHC1991}, it becomes difficult to build and maintain a dictionary for the vocabulary that appears in each text.
    Therefore, methods for tagging syllable units without a dictionary have been proposed~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016} and studies have been conducted to improve them ~\cite{KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    From a mechanical perspective, syllable-by-syllable morphological analysis can be performed either by tagging syllable-by-syllable and then applying a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016}, or by tagging syllable-by-syllable with the base form already restored~\cite{YounJY2021}.
    However, syllable-by-syllable morphological analysis has limitations, in that it is difficult to accurately identify morpheme boundaries and learn long-term contextual information as the length of the sequence increases.
    In this study, the former is referred to as dictionary-based morphological analysis and the latter as syllable-unit morphological analysis.
    Both methods are trained on a manually labeled corpus and cannot accurately analyze new syllable combinations or morphemes that do not appear in the training corpus.
    Recently, with the development of the Internet and the spread of open sources as well as open data, web texts, corpora, language resources, and knowledge shared by different people have accumulated significantly.
    The reduced cost of building and maintaining a dictionary provides a significant opportunity to overcome the limitations of dictionary-based methods.

    Against this background, this study considers how the dictionary-based morphological analysis method used by MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis in tokenizers, which is an essential preprocessing tool for deep learning, can be effectively improved, and a method is proposed.
    The dictionary-based morphological analysis method~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018} trained by the CRF (Conditional Random Fields) method~\cite{Lafferty2001} lists the candidate morphemes in the dictionary from a given sentence to form a lattice structure connected by a directed graph and determines the optimal morphological analysis path within it.
    The process of determining the optimal path in the lattice uses the Viterbi algorithm~\cite{Viterbi1967}, which determines the path that minimizes the cost of each morpheme node and the sum of the neighborhood costs of two consecutive morphemes.
    The main types of errors in these dictionary-based morphological analysis methods occur when new words that are not in the dictionary are used in a sentence, or when the optimal path calculation selects the incorrect result owing to bias.
    For example, it may be cost-effective to select one long morpheme than several short morphemes, but this may often lead to an incorrect analysis.
    The main motivation for this study was that the path that minimizes the costs for the nodes and links may not be the optimal path.

    To identify cases in which a suboptimal solution is actually the best solution according to the best path calculation, we modified the best path calculation method to generate suboptimal analysis results and verified the extent to which they are correct.
    Although there are numerous different approaches to select the next-best path, we used the method of replacing a morpheme node on the optimal path with a lower-ranked node.
    As shown in Table~\ref{tab:maximum-performance}, we confirm the extent to which the analysis performance can be improved by replacing the optimal path with a lower-ranked node.
    We can consider the problem of finding the correct answer among the generated sub-optimals, similar to the problem of re-ranking search results in information retrieval~\cite{BaeYJ2021}.
    In ~\cite{ChoiYS2018}, the N-best analysis results generated by the seq2seq model were re-ranked based on a convolutional neural network to improve the performance.
    In this study, re-ranking was performed using two BERT models of different types and forms, as proposed in ~\cite{Nogueira2019}.
    Experimental results show that first-stage re-ranking improves the performance by over 20\% over previously written and spoken models, and second-stage re-ranking with a different type of input and a different type of pre-trained model further improves the performance by more than 30\% over previously written and spoken models.

    With this method, the performance of the dictionary-based morphological analysis method could be further improved; however, the overall analysis time increased when the morphological analysis system was configured, including the re-ranking model itself.
    However, it is feasible to use the results of multiple reranked morpheme analyses to update the connection costs between morphemes in a dictionary, similar to the backpropagation process in a typical neural network.
    It is also expected that the morphological analysis system with improved connection costs will be able to generate better re-ranking candidates, which will further improve performance by doing so iteratively.
    Further research is required in the future.
    In this study, only performance improvement using the second-stage re-ranking was covered as the scope of the study.
    The main contributions of this study are as follows.
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We explore the possibility of performance improvement by introducing a method to replace the optimal path with a suboptimal node and propose a method to effectively improve the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To improve the performance of dictionary-based analysis by re-ranking the morphological analysis results, we propose extending the performance improvement using different BERT models to perform two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We propose a new method for updating dictionary connection costs based on re-ranked morphological analysis results. We also outline directions for future research, suggesting potential improvements.
    \end{enumerate}
    These contributions provide important insights into the performance improvement of Korean morphological analysis and the direction of future research, and will serve as a useful reference for future researchers.

    The remainder of this paper is organized as follows.
    In Section ~\ref{sec:related-work}, we introduce previous research cases related to this study. In Section ~\ref{sec:morphological-analysis-model}, we discuss configuring and training a dictionary-based morphological analysis system.
    In Section ~\ref{sec:re-ranking-model}, we discuss the generation of secondary results of morphological analysis, produce re-ranking data, and propose a method for training a two-stage re-ranking model.
    In Section ~\ref{sec:results}, we discuss the results of the performance improvement using the morphological analysis and re-ranking models.
    Finally, in Section ~\ref{sec:conclusion}, we conclude the study and discuss its limitations and directions for future research.

@@

    Korean morphological analyses have seen an influx of various methodologies in recent years~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}. The nature of the Korean language, being agglutinative, introduces challenges that have propelled researchers to develop inventive solutions, many of which have laid the groundwork for future research.

    \subsection{Traditional Dictionary-based Approaches}\label{subsec2.1}
    The earliest attempts at Korean morphological analysis relied heavily on rule- and dictionary-based methods~\cite{KwonHC1991}. These methodologies employ predefined sets of linguistic rules or large dictionaries to detect morphemes and determine parts of speech. One of the most significant advantages of this approach is its deterministic nature, which can lead to high accuracy when the input text closely adheres to the rules or dictionaries used. However, it is challenging to scale and update them, particularly with the constant evolution of language and the emergence of new words. The dynamic nature of language, particularly in the Internet age, has made the maintenance of comprehensive dictionaries labor-intensive.

    \subsection{Syllable-unit Morphological Analysis}\label{subsec2.2}
    To overcome the limitations of dictionary dependence, syllable-by-syllable morphological analysis has emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}. This method either tags each syllable and then applies a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}. One notable drawback is the challenge of accurately pinpointing morpheme boundaries. Furthermore, as the sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Deep Learning and Tokenization}\label{subsec2.3}
    The advent of deep learning has reshaped the Korean morphological analysis landscape. High-performance deep learning models in NLP employ tokenization, dividing the text into smaller units and then converting these tokens into vectors as an input to the computational model~\cite{Mikolov2013}. The attempt to conduct subword tokenization, incorporating separate morphemes to cater to Korean linguistic characteristics is particularly significant. Incorporating accurate morpheme analysis results into tokenization can improve overall analysis performance.

    With advancements in technology and access to large datasets, the potential to integrate dictionary-based methods with deep learning techniques has emerged. As motivated in our study, MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis, utilized dictionary-based morphological analysis trained by the CRF method~\cite{Lafferty2001}. This method attempts to find the optimal morphological analysis path by forming a lattice structure as in ~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018}. The amalgamation of dictionary-based methods with deep learning, as explored in our research, signifies the latest strides in this journey aimed at harnessing the strengths of both approaches. Future research avenues include refining these hybrid methods and exploring the capabilities of end-to-end models.

@@

    A representative corpus for training Korean morphological analysis models is the Sejong morphological analysis corpus, which is often referred to as the Sejong corpus. The Sejong corpus, a major output of the 21st Century Sejong Project, contains 15 million words and is the mainstay of Korean morphological analysis research~\cite{ChoeMW2008}.

    Subsequent efforts have extended and refined this foundational work. The University of Ulsan addressed the limitations of the Sejong corpus~\cite{KimIH2010} and introduced [UCorpus], an expanded corpus of 63 million words. This expanded corpus not only adds new data but also corrects errors in the original Sejong corpus~\cite{UCorpusHG}. To further extend this step, the National Institute of Korean Language launched a massive Korean corpus construction project in 2018. The fruits of this labor, a corpus called <Everyone Corpus>, were released to the public in 2020~\cite{EveryoneCorpus}. A special offshoot of this project produced a morpheme-tagged corpus of three million words, including spoken languages ~\cite{KimIH2019}.

@@

    To train a dictionary-based morpheme analysis model effectively, the morpheme-tagged corpus, typically represented in lemma form, must be transformed to include the boundary information between morphemes in its surface form. Crucial to this transformation is string alignment, a process that accounts for the discrepancies between lemma forms and surface forms in the Korean morphological analysis corpus.

    In this study, string alignment was performed using the Smith--Waterman algorithm, which uses a scoring matrix based on the similarity of the grapheme unit of Korean letters for each word pair (as shown in Figure~\ref{fig:sample}). Each aligned sentence containing a morpheme tag was converted into a training sample tailored for dictionary-based morphological analysis.

    The resulting table in Figure~\ref{fig:sample} illustrates this process. Each row acts as a lexical unit. The first four columns contribute to feature generation, whereas the last four columns facilitate post-lemmatization. Using the morphological corpus above, a large number of training samples can be generated according to the process shown in Figure~\ref{fig:sample}. With the exception of the evaluation samples, the remaining sentences were used to train the dictionary-based morphological analysis model using the CRF algorithm. The output of this training allowed the calculation of the costs associated with each morpheme node and the linking of two consecutive morphemes. This, in turn, allowed the discovery of an optimal path using the Viterbi algorithm.

@@

    Figure~\ref{fig:lattice} presents a snapshot of the lattice structure, which is an integral part of the morphological analysis. (1) shows a fragment of the lattice structure formed when the example sentence in Figure~\ref{fig:sample} is entered. (2) shows the optimal path determined using the Viterbi algorithm.

    However, the path inferred by the trained model may differ from the correct solution constructed by humans. The nodes marked with stars in (1) represent correct nodes. The upper-left number of each node indicates the ranking of the nodes accessible at each decoding point. The choices made at certain moments deviate from the correct solution. To improve analytical performance, mechanisms to correct these discrepancies must developed.

@@

    Although dictionary-based morphological analysis offers significant advances, its optimal paths occasionally diverge from the correct solutions that humans understand. This divergence underscores the need for a model that reevaluates these primary results and reorients them to achieve higher accuracy. This method, called "re-ranking," involves generating multiple analyses of an input and then reordering them based on a new set of criteria or models, thereby improving the overall quality of the results.

@@

    Before reranking begins, multiple analyses, typically referred to as the N-best paths, of the input sentence are generated. This involves extracting the top N atoms from the lattice structure. In this study, a novel approach is introduced to generate secondary paths, as shown in (3) of Figure~\ref{fig:lattice}, by selecting the second-best node rather than each best node constituting the path from the best-path result. Some of these secondary paths provided alternatives that reconciled incorrect with correct answers. Similarly, paths modified by favoring the third-best node were called tertiary paths, and this naming convention was continued for subsequent paths. In our preliminary test, the secondary paths, including the optimal and suboptimal paths, were shown to cover the majority of the correct morphological analyses as measured through human evaluations. (Refer to Table~\ref{tab:maximum-performance}).

@@

    Bidirectional Encoder Representations from Transformers (BERT) models~\cite{Devlin2019} have revolutionized many natural language processing tasks by understanding the context in which words appear in text. In this study, we attempt to leverage the power of BERT to reorder the generated secondary paths. We labeled the generated secondary paths with scores related to the morphological analysis performance to fine-tune a pre-trained BERT model specialized for Koreans with excess amount Korean text. After pre-testing several scoring methods on a modest scale, we found that using scores based on the degree of error rather than scores based on accuracy by widen the gap between correct and incorrect answers is effective for learning.

    Once the BERT model is fine-tuned and trained for the re-ranking task, it can predict a re-ranking score for each path in the secondary path list. This implies that considering the context, morphological organization, and other essential linguistic features of the path, the model assigns a score to each path. The paths were then re-ranked according to this score, and the path with the highest score was selected for the best morphological analysis.

@@

    Given the complexity of the Korean language, a single re-ranking step does not constantly yield accurate results. Therefore, we propose a two-step re-ranking approach as described in ~\cite{Nogueira2019}.

    In the first step, we re-rank the secondary paths generated using the BERT model, as described in Section ~\ref{subsec4.3}. In the second step, we introduced another BERT variant that was optimized for a different set of linguistic features or trained on a different dataset. This allowed us to perform a fine-grained re-evaluation, further refine the list, and push more contextually accurate paths to the top.

    As shown in Figure~\ref{fig:ranking}, for a two-stage reranking model, the first stage performs the first re-ranking, taking as input a secondary path in morphologically tagged lemma form. It then performs a second re-ranking, again taking as input the path re-ranked in stage 1 and the original input sentence. This was conducted to improve effectiveness, as it was ineffective given the same type of input.

@@

    We evaluated the performance of the proposed deep learning-integrated dictionary-based morphological analysis method. This section presents the results of the experimental evaluation considering the improvements over conventional methods and the effectiveness of our re-ranking model.

@@

    We found that the proposed transformer-based re-ranking technique consistently improved the results of the existing morphological analysis models. These results open new possibilities in the area of Korean morphological analysis by further improving the results of existing traditional machine learning models.
    Finally, because the results of the major related works proposed in the literature were mostly conducted on the Sejong corpus, we compared the performance improvement of the Sejong corpus with the results of previous studies.
